---
title: "VPRMLandSurfaceModel"
author: "Timothy W. Hilton"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction to VPRMLandSfcModel.

The VPRMLandSfcModel R package provides a pure R implementation of the
Vegetation Photosynthesis and Respiration Model (VPRM)
of [Mahadevan et al. (2008)](https://doi.org/10.1029/2006GB002735), along with tools for assembling VPRM
driver data and estimating model parameter values.

The following examples illustrate how to use the VPRMLandSfcModel R
package.  The documentation for individual functions explains their
purpose, output, and arguments in much more detail.

## Usage Example: assembling VPRM driver data

VPRMLandSfcModel can accept input data as either an R data frame or as
an object of class VPRM_driver_data.  Using the included dataset
`ParkFalls`, run the VPRM for the [Park Falls, Wisconsin, USA Ameriflux
site](http://ameriflux.lbl.gov/sites/siteinfo/US-PFa):
```{r, fig.show='hold'}
library(VPRMLandSfcModel)
data(Park_Falls)
## determine the phenology phase for each tower observation date
phen_filled <- interp_phenology(PFa_phen, PFa_tower_obs[['date']])
## place the tower observations and MODIS data into a VPRM_driver_data object.
pfa_dd <- VPRM_driver_data(name_long="Park Falls",
                           name_short = "US-PFa",
                           lat=45.9459,
                           lon=-90.2723,
                           PFT='MF',    ## mixed forest
                           tower_date=PFa_tower_obs[['date']],
                           NEE_obs=PFa_tower_obs[['FC']],
                           T=PFa_tower_obs[['TA']],
                           PAR=PFa_tower_obs[['PAR']],
                           date_nir = PFa_refl[['date']],
                           rho_nir=PFa_refl[['nir']],
                           date_swir = PFa_refl[['date']],
                           rho_swir = PFa_refl[['swir']],
                           date_EVI = PFa_evi[['date']],
                           EVI=PFa_evi[['evi']],
                           phen=phen_filled)
## take a look at the result
print(head(as.data.frame(pfa_dd)))
```

Now plot air temperature.  The [ggplot2](http://ggplot2.tidyverse.org) package is not necessary
to install or use VPRMLandSfcModel, but it makes very pretty plots.

```{r, fig.show='hold'}
library(ggplot2)
library(chron)
##
fig_T <- (ggplot(pfa_dd[['data']],
              aes(date, T)) +
        geom_line() +
        scale_x_chron(format="%d %b %Y") +
        labs(title = "US-PFa", y=expression(air~T~(degree*C))) +
        theme_classic() +
        theme(plot.title = element_text(hjust = 0.5)) +    # center the plot title
		theme(axis.text=element_text(size=14),
			axis.title=element_text(size=14,face="bold")))
print(fig_T)
```

## Usage Example: calculating VPRM NEE

Calculating VPRM fluxes requires values for the parameters lambda,
PAR$_0$, alpha, and beta.  The dataset VPRM_parameters, part of the
VPRMLandSfcModel package, includes several parameters sets.

```{r, fig.show='hold'}
data(VPRM_parameters)
attach(all_all_VPRM_parameters)
pfa_dd[['data']][['VPRM_NEE']] <- vprm_calc_NEE(
    pfa_dd, lambda=lambda, PAR_0=PAR_0, alpha=alpha, beta=beta)

fig_NEE_VPRM <- (ggplot(pfa_dd[['data']],
	aes(date, VPRM_NEE)) +
	geom_point() +
	scale_x_chron(format="%d %b %Y") +
	labs(title = "US-PFa", x='date', y=expression(VPRM~NEE~(mu*mol~m^{-2}~s^{-1}))) +
	theme_classic() +
	theme(plot.title = element_text(hjust = 0.5)) +    # center the plot title
	theme(axis.text=element_text(size=14),
		axis.title=element_text(size=14,face="bold")))
print(fig_NEE_VPRM)

## now plot the eddy covariance-observed NEE
fig_NEE_EC <- (ggplot(pfa_dd[['data']],
	aes(date, NEE_obs)) +
	geom_point() +
	scale_x_chron(format="%d %b %Y") +
	labs(title = "US-PFa", x='date',
		y=expression(eddy~covariance~NEE~(mu*mol~m^{-2}~s^{-1}))) +
	theme_classic() +
	theme(plot.title = element_text(hjust = 0.5)) +    # center the plot title
	theme(axis.text=element_text(size=14),
		axis.title=element_text(size=14,face="bold")))
print(fig_NEE_EC)

## now plot the difference between covariance-observed NEE and VPRM NEE
fig_dNEE <- (ggplot(pfa_dd[['data']],
	aes(date, NEE_obs-VPRM_NEE)) +
	geom_point() +
	scale_x_chron(format="%d %b %Y") +
	labs(title = "US-PFa", x='date',
		y=expression(Delta*NEE[obs-VPRM]~(mu*mol~m^{-2}~s^{-1}))) +
	theme_classic() +
	theme(plot.title = element_text(hjust = 0.5)) +    # center the plot title
	theme(axis.text=element_text(size=14),
		axis.title=element_text(size=14,face="bold")))
print(fig_dNEE)
```

The difference between the eddy covariance NEE and the VPRM NEE
illustrates that the all-all VPRM parameter set
from [Hilton et al. (2013)](https://www.biogeosciences.net/10/4607/2013/) isn't the best choice (except for ease
of use).  [Hilton et al (2014)](https://www.biogeosciences.net/11/217/2014/) discusses different spatial and
temporal groupings of data for parameter estimation and concludes that
grouping by plant functional types, perhaps in annual windows,
minimizes model error among the options considered.

# Usage Example: estimating VPRM parameters #

The following code estimates two sets of parameter values for
the [Park Falls, Wisconsin Ameriflux site](http://ameriflux.lbl.gov/sites/siteinfo/US-PFa) using the methodology
described in [Hilton et al. (2013)](https://www.biogeosciences.net/10/4607/2013/).  The first set estimates
values for the entire year. The second estimates monthly values,
creating 12 sets of values for [lamba, PAR_0, alpha, beta].

The examples below set DE_itermax to a very low value (2) so that the
examples run quickly.  To obtain useful parameter estimates,
DE_itermax must be set much larger.  500 to 1000 is usually a good
starting point.

```{r, fig.show='hold'}
library(VPRMLandSfcModel)
library(chron)

data(Park_Falls)
pfa_dd <- VPRM_driver_data(name_long="Park Falls",
                           name_short = "US-PFa",
                           lat=45.9459,
                           lon=-90.2723,
                           PFT='MF',
                           tower_date=PFa_tower_obs[['date']],
                           NEE_obs=PFa_tower_obs[['FC']],
                           T=PFa_tower_obs[['TA']],
                           PAR=PFa_tower_obs[['PAR']],
                           date_nir = PFa_refl[['date']],
                           rho_nir=PFa_refl[['nir']],
                           date_swir = PFa_refl[['date']],
                           rho_swir = PFa_refl[['swir']],
                           date_EVI = PFa_evi[['date']],
                           EVI=PFa_evi[['evi']],
                           phen=NA)

## estimate parameter values for all data
par_est_status <- estimate_VPRM_pars(all_data=pfa_dd[['data']],
                                     DE_itermax = 2,
                                     par_set_str='ExampleRun')

## estimate parameter values for monthly windows
par_est_status <-
    estimate_VPRM_pars(all_data=pfa_dd[['data']],
                       DE_itermax = 2,
					   out_path=,
                       par_set_str='ExampleRun_Monthly',
                       opt_groups=list(months(pfa_dd[['data']][['date']]),
                                       years(pfa_dd[['data']][['date']])))
```

`estimate_VPRM_pars` writes the best-fit parameter values to an RData
file in the working directory.  After the parameter estimation
completes, load the RData file into the workspace to have a look.  The
optimized parameters are a collection of objects of class DEoptim.
See the documentation for the [DEoptim package](https://cran.r-project.org/package=DEoptim) for details.  The
best-fit values are in the field optim$bestmem.
